Mental Health Screening in CF

### 2023-08-19 Emily Muther/mental_health_screening.qmd
cf mental health ZINB not working

@@

```{r include=FALSE}
library(tidyverse)
library(plotly)
library(gtsummary)
library(GLMMadaptive)
# Import
df = read.csv("/Volumes/PEDS/RI Biostatistics Core/Shared/Shared Projects/Vigers/CF/Emily Muther/Mental Health Screening/Data_Cleaned/MH master data.csv",check.names = F,na.strings = "")
# Column types
df$CONTACT_DATE = lubridate::mdy(df$CONTACT_DATE)
df$`PHQ-9 Total Score (calculated)` = as.integer(df$`PHQ-9 Total Score (calculated)`)
# Visits
df = df %>% arrange(MRN,CONTACT_DATE) %>% group_by(MRN) %>%
  mutate(`Visit Number` = row_number(), 
         `Days From Baseline` = as.numeric(CONTACT_DATE - CONTACT_DATE[1]))
```

@@

# Methods

Because the PHQ-9 results are right-skewed with lots of 0 scores, we used a zero inflated Poisson mixed model to analyze change in score over time.

>>>>

```{r}
#| include: false
library(tidyverse)
library(plotly)
library(gtsummary)
library(GLMMadaptive)
library(lmerTest)
library(broom.mixed)
library(knitr)
library(sjPlot)
library(report)
```

@@

```{r}
#| warning: false
# Import
df = read.csv("/Users/timvigers/Library/CloudStorage/Dropbox/Work/Vigers/CF/Emily Muther/Mental Health Screening/Data_Cleaned/MH master data.csv",check.names = F,na.strings = "")
# Column types
df$MRN = factor(df$MRN)
df$CONTACT_DATE = lubridate::mdy(df$CONTACT_DATE)
df$`PHQ-9 Total Score (calculated)` = as.integer(df$`PHQ-9 Total Score (calculated)`)
df$`GAD-7 Total Score` = as.integer(df$`GAD-7 Total Score`)
# Visits
df = df %>% arrange(MRN,CONTACT_DATE) %>% group_by(MRN) %>%
  mutate(`Visit Number` = row_number(), 
         `Days From Baseline` = as.numeric(CONTACT_DATE - CONTACT_DATE[1]))
```

@@

# Results

## ZINB mixed effects model

Because the PHQ-9 results are right-skewed with lots of 0 scores, we first tried a zero inflated negative binomial (ZINB) mixed model to analyze change in score over time. At first we tried using categorical visit number as a time variable, but the model did not converge (likely due to the number of parameters).

```{r}
#| warning: false
try(mixed_model(`PHQ-9 Total Score (calculated)` ~ factor(`Visit Number`),random = ~1|MRN,data = df,family = zi.negative.binomial(), zi_fixed = ~ 1))

try(mixed_model(`PHQ-9 Total Score (calculated)` ~ `Days From Baseline`,
                 random = ~1|MRN,data = df,family = zi.negative.binomial(), 
                 zi_fixed = ~ 1))
```

Because the ZINB models were not converging, we instead tried to use a simpler Poisson mixed model (without accounting for the inflated 0s) with random intercept per participant. Visit number was treated as a continuous variable because this model would not converge with categorical visit numbers.

```{r}
#| warning: false
m2 = glmer(`PHQ-9 Total Score (calculated)` ~ `Visit Number` + (1|MRN),data = df,family = "poisson")
plot_model(m2,type="pred",
           terms=c("Visit Number","MRN"),
           pred.type="re", ci.lvl = NA) + 
  theme(legend.position = "none")
res = tidy(m2,exponentiate = T,effects="fixed")
kable(res,digits = 3)
```

## Simpler models

Because the complex models were difficult to fit and not particularly informative, we instead modeled the odds of an elevated score for the full cohort, and the change in score over time for those with elevated scores. Estimates are reported as odds ratios in tables, but are reported on the log scale in the text, just in case both are needed.

### PHQ9

```{r}
df$PHQ9_elevated = factor(df$`PHQ-9 Total Score (calculated)` >= 5,
                          levels = c(F,T),labels = c("No","Yes"))
```

A PHQ9 score $\geq$ 5 was considered elevated.

#### Odds of an elevated score

```{r}
mod = glmer(PHQ9_elevated ~ `Visit Number` + (1|MRN),data = df,family = "binomial")
res = tidy(mod,exponentiate = T,effects="fixed")
kable(res,digits = 3)
```

`r report(mod)`

#### Association with visit number in those with elevated scores.

Results are reported as expected counts.

```{r}
mod = glmer(`PHQ-9 Total Score (calculated)`~ `Visit Number` + (1|MRN),
            data = df[df$PHQ9_elevated=="Yes",],family = "poisson")
res = tidy(mod,exponentiate = T,effects="fixed")
kable(res,digits = 3)
```

`r report(mod)`

### GAD7

```{r}
df$GAD7_elevated = factor(df$`GAD-7 Total Score` >= 4,
                          levels = c(F,T),labels = c("No","Yes"))
```

A GAD7 score $\geq$ 4 was considered elevated.

#### Odds of an elevated score

```{r}
mod = glmer(GAD7_elevated ~ `Visit Number` + (1|MRN),data = df,family = "binomial")
res = tidy(mod,exponentiate = T,effects="fixed")
kable(res,digits = 3)
```

`r report(mod)`

#### Association with visit number in those with elevated scores.

Results are reported as expected counts.

```{r}
mod = glmer(`GAD-7 Total Score`~ `Visit Number` + (1|MRN),
            data = df[df$GAD7_elevated=="Yes",],family = "poisson")
res = tidy(mod,exponentiate = T,effects="fixed")
kable(res,digits = 3)
```

`r report(mod)`
